---
title: "Project Assignment 1"
author: "muya"
date: "5/29/2022"
output: html_document
---

Loading and preprocessing the data

```{r}
activityData <- read.csv(file="activity.csv", header=TRUE)
```

Obtaining the mean total number of steps taken each day

```{r}

# Get the total steps taken each day
total_Steps <- aggregate(steps ~ date, activityData, FUN=sum)

# Make a histogram of the total number of steps taken each day
hist(total_Steps$steps,
     main = "Total Steps each Day",
     xlab = "Number of Steps")
```

```{r}
# Obtain the mean and median of total steps taken each day
mean_Steps <- mean(total_Steps$steps, na.rm = TRUE)
median_Steps <- median(total_Steps$steps, na.rm = TRUE)
```
Mean Number of Steps Taken each Day = 1.076618910^{4}    
Median Number of Steps Taken each Day = 10765

What is the average daily activity pattern?

```{r}
# Make a time-series plot of the 5-minute interval and the average number of
# steps taken, averaged across all days.
library(ggplot2)
mean_Steps_By_Int <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = mean_Steps_By_Int, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("The Average Daily Activity Pattern") +
  xlab("five minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
# Which five minute interval across all days contain the maximum number of steps
max_Int <- mean_Steps_By_Int[which.max(mean_Steps_By_Int$steps),]
```


# Imputing Missing Values

```{r}
# Calculate and report the total number of missing values in the dataset
missing_Values <- is.na(activityData$steps)

# Devise a strategy for filling in all of the missing values
```
There are 17568 missing values. I will replace these missing values with the five day average of that respective interval.

```{r}
# Create a new dataset that is equal to the original dataset but with the missing data filled in.

imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             mean_Steps_By_Int$steps[match(activityData$interval, 
                                                                        mean_Steps_By_Int$interval)],
                                             activityData$steps))


```
```{r}
# Make a histogram of the total number of steps taken each day and
# and report the mean and median.
impute_Steps_By_Int <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impute_Steps_By_Int$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")
```

```{r}
impute_Mean_Steps <- mean(impute_Steps_By_Int$steps, na.rm = TRUE)
impute_Median_Steps <- median(impute_Steps_By_Int$steps, na.rm = TRUE)
diff_Mean = impute_Mean_Steps - mean_Steps
diff_Median = impute_Median_Steps - median_Steps
diff_Total = sum(impute_Steps_By_Int$steps) - sum(total_Steps$steps)
```
There is a difference of 0 in the mean steps of the two dataset. There is a difference of -1.076381110^{4} in the median steps of the two dataset.    There is a difference of 8.612950910^{4} in the total steps of the two dataset.

# Are there differences in activity patterns between weekdays and weekends?

```{r}

# Create a new factor variable in the dataset with two levels - "weekend" and "weekday"
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
      return ("weekend")
   
 
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)

```


